### 2025-03-28 paper.tex
Edit design considerations

% TODO: Figure out how to call those "programming experiences" (also is this a good name?)
%   the text uses "concrete, collaborative, interactive" - but this is a silly list
%   really, they are "cool" but how to say that.
%
% TODO: Also, what exactly goes into the list?
%
%
\documentclass[sigconf,anonymous,screen]{acmart}
%\documentclass[sigconf,review,anonymous]{acmart}

\usepackage{hhline,xcolor,colortbl}
\usepackage{fontawesome}
\usepackage{enumitem}
\usepackage{fancyvrb}
\usepackage{tikz}
\usepackage[breakable,theorems,skins]{tcolorbox}
\usepackage{lettrine}
\usepackage{hyperref}

\newcommand{\ident}[1]{{\sffamily #1}}
\newcommand{\srcid}[1]{\textnormal{\sffamily #1}}
\newcommand{\srckvd}[1]{\textnormal{\sffamily\bfseries #1}}
\newcommand{\note}[1]{\textcolor{red}{#1}}

\newcommand*\circled[1]{\textnormal{\footnotesize\sffamily\bfseries\protect\tikz[baseline=(char.base)]{
  \node[shape=circle,fill=black,text=white,draw,inner sep=1pt] (char) {#1};}}}

\DeclareRobustCommand{\keyideabox}[3]{\begin{tcolorbox}[breakable,
  boxsep=5pt,left=0pt,right=0pt,top=0pt,bottom=0pt,width=\dimexpr\columnwidth\relax,
  colback=gray!20,colframe=gray!20,
  enlarge bottom by=0pt,enlarge top by=0pt,
  arc=0pt,outer arc=0pt]
\lettrine[lraise=0.3]{\LARGE #1}{~}
\small \textbf{#2.} #3
\end{tcolorbox}
}

\setlist{leftmargin=1.5em}
\setlength\itemsep{0.5em}

\setcopyright{none}
\copyrightyear{2025}
\acmYear{2025}
\acmDOI{XXXX.XXXX}

\acmConference[Conference acronym 'XX]{}{June 03--05, 2018}{Woodstock, NY}
\acmISBN{978-1-4503-XXXX-X/18/06}

\begin{document}
\title[Denicek: Computational Substrate for Document-Oriented End-User
  Programming]{{\scshape Denicek}: Computational Substrate for\\ Document-Oriented End-User Programming}

\author{Tomas Petricek}
\email{tomas@tomasp.net}
\orcid{0000-0002-7242-2208}
\affiliation{%
  \institution{Faculty of Mathematics and Physics, Charles University}
  \city{Prague}
  \country{Czech Republic}
}

\begin{abstract}
User-centric programming research gave rise to a variety of compelling programming experiences,
including local-first collaborative editing, programming by demonstration, incremental
recomputation, schema change control, end-user debugging and concrete programming.
Those experiences advance the state of the art of end-user programming, but they are hard to
implement on the basis of established programming languages and system.

We contribute Denicek, a computational substrate that simplifies the implementation of the
above programming experiences. Denicek represents a program as a series of edits
that construct and transform a document consisting of data and formulas. It provides three
operations on edit histories: edit application, merging of histories and conflict resolution.
By composing these operations, we can  easily implement a range of programming experiences.

We present the architecture of Denicek, discuss key design considerations and elaborate
the implementation of a variety of programming experiences. To evaluate the proposed
substrate, we use Denicek to develop an innovative interactive data science notebook system.
The case study shows that the Denicek computational substrate provides a suitable basis for the
design of rich, interactive end-user programming systems.
\end{abstract}

\keywords{Programming Systems, Computational Substrate, End-User Programming,
 Programming by Demonstration, Local-First Software}

%\begin{teaserfigure}
%  \includegraphics[width=\textwidth]{sampleteaser}
%  \caption{Seattle Mariners at Spring Training, 2010.}
%  \Description{Enjoying the baseball game from the third-base
%  seats. Ichiro Suzuki preparing to bat.}
%  \label{fig:teaser}
%\end{teaserfigure}
\maketitle

% ==================================================================================================

\section{Introduction}

A computational substrate defines the structures with which programs are constructed, how
the progam state is represented and how it evolves \cite{jakubovic-2022-ladder}. The choice of
a substrate affects what programming experiences can be readily supported by a system. For example,
object-oriented programming has been historically linked to graphical user
interfaces~\cite{kay-1993-smalltalk}, while representing programs as lists enabled Lisp to become
a language laboratory \cite{steele-1993-lisp}.

In principle, any programming experience can be developed on top of any computational
substrate. However, a suitable program representation can eliminate much of the complexity of implementing interesting
programming experiences. For example, the reflective capabilities of Smalltalk make it easy
to build rich debugging tools \cite{rauch-2019-babylonian} that are difficult to implement
for C/C++~\cite{kell-2018-unix,kell-2024-debugging}.

\subsubsection*{Programming Experiences}

In this paper, we describe a programming substrate that makes it easy to develop diverse
programming experiences \cite{myers-2006-eup}. Those include:

\begin{itemize}
\item \emph{Local-First Collaboration.} Multiple users should be able to concurrently
  modify a single document and automatically merge their changes, preferrably without a central server~\cite{kleppmann-2019-local,litt-2022-peritext}.
\item \emph{Programming by Demonstration.} Allow users to construct simple programs by
  enacting the steps of the expected behavior using concrete examples~\cite{leiva-2021-rapido,cypher-1993-pbd}.
\item \emph{Incremental Recomputation.} When a part of a document changes, formulas whose result
  depends on it are invalidated and, possibly, automatically recomputed \cite{teitelbaum-1981-cps,mcdirmid-2013-usable,horowitz-2023-engraft}.
\item \emph{Schema Change Control.} When the user evolves the structure of the document, affected
  data and formulas should automatically co-evolve to match the new structure~\cite{litt-2020-cambria,edwards-2025-schema}.
\item \emph{End-User Debugging.} The user should be able to ask provenance questions~\cite{cheney-2009-provenance} --
  why a computation resulted in a particular value and what inputs contributed to the result \cite{ko-2009-whyline}.
\item \emph{Concrete Programming.} It should be possible to reuse parts of program logic, or formulas,
  without introducing abstractions, that is, program against concrete values \cite{edwards-2006-copypaste,edwards-2022-copypaste}.
\end{itemize}

\subsubsection*{Two-Stage Methodology}
The technical focus of this paper fits within the interior mode of design science research \cite{adam-2021-dsr}.
To \emph{design} Denicek, we identify six \emph{formative examples} -- simple programming
tasks that manifest one or more of the desired programming experiences~(\S\ref{app:examples}).
Using those examples, we co-design the Denicek substrate and Webnicek~(\S\ref{sec:walk}),
a simple web-based end-user programming environment built directly on top of the substrate. Although Webnicek can
be used to complete end-user programming tasks, it is optimized for developing the underlying
substrate rather than for usability.

To \emph{evaluate} Denicek, we use the substrate to develop Datnicek (\S\ref{sec:case}), an interactive data
science notebook system inspired by existing environments~\cite{kandel-2011-wrangler,drossos-2020-wrex}.
We use the second stage to evaluate suitability of the Denicek substrate for the development of
end-user programming systems and report the results of our evaluation (\S\ref{sec:eval}).

\subsubsection*{Substrate}
The Denicek substrate brings together two central design ideas. First, it represents programs as
document trees consisting of nodes that can represent data, formulas, evaluated results, as well as
static content. Second, Denicek does not store the document tree itself, but instead, maintains
a sequence of edit operations through which the tree was constructed and transformed.

The substrate then provides three primitive operations for working with sequences of edits.
First, it can apply a series of edits to reconstruct the document. Second, it can merge two
diverging edit histories. Finally, it can detect conflicts when merging histories and, for
example, remove conflicting edits from one branch.

The key insight that we present in this paper is that many compelling programming experiences can
be implemented by leveraging the uniform document representation alongside with a suitable
composition of the three primitive operations.

Changing data and formulas is done using the same primitive edit operations that
manipulate the document structure. A user-interface may provide a specialized editor, but still
trigger the primitive edits behind the scenes. However, interacting with elements in the document,
such as a entering text in a textbox can also generate a document edit that can be merged or checked
for conflicts (\S\ref{sec:impl-interaction}).

As we will see, past edits that demonstrate an operation done to the
document can be recorded, allowing programming by demonstration (\S\ref{sec:impl-pbd}).  Replaying
such recorded edits is implemented using the merging operation, which means that recorded operations
continue working even if the document structure later evolves. Moreover, structural changes to the
document can be merged with concurrent data edits  (\S\ref{sec:impl-collab}). Evaluation of formulas
also generates document edits (\S\ref{sec:impl-eval}). If the evaluated edits conflict with manual
edits done later by the user, the evaluated edits are removed, implementing an incremental
recomputation (\S\ref{sec:impl-incremental}).

\subsubsection*{Contributions}
The structure and contributions of this paper are:

\begin{itemize}
\item We present the Denicek substrate (\S\ref{sec:system}) and provide a detailed description of its
  document representation, edit operations and the key three operations for working with edit histories.
\item We illustrate a range of end-user programming experiences supported in Webnicek, a simple
  web-based prototype programming system (\S\ref{sec:walk}), and discuss how the experiences are
  implemented using the Denicek substrate (\S\ref{sec:impl}).
\item We document important design decisions, alternatives and limitations (\S\ref{sec:discuss}).
  The analysis shows that the desired functionality requires a careful choice among interconnected
  design options.
\item To evaluate how Denicek simplifies the development of program\-ming systems, we
  build an innovative data science notebook Datnicek (\S\ref{sec:case})
  and assess its impementation complexity~(\S\ref{sec:eval}).
\end{itemize}

\noindent
To enable others build on top of Denicek, we share our compact and documented
source code at: \url{https://github.com/removed/for/review} %d3sprog/denicek

% ==================================================================================================

\section{Background}
\label{sec:background}
Denicek aims to support a class of systems associated with end-user programming \cite{myers-2006-eup},
notational freedom and self-sustainability~\cite{jakubovic-2023-techdims}, liveness, interactivity
and richness \cite{horowitz-2023-qualities,rein-2019-live}. We see programming more as interacting with a medium
or a substrate \cite{kay-1977-media,klokmose-2015-webstrates} than writing text \cite{rpg-2012-revolution}.
We focus on systems that allow gradual progression from a user to a developer, as envisioned
in Smalltalk~\cite{reenskaug-1981-byte} and use the term end-user programming loosely to refer to a
part of this spectrum, also including spreadsheet systems and notebooks for data science.
We first review related systems and their technical design (\S\ref{sec:background-sys})
before discussing programming experiences (\S\ref{sec:background-exp}).

% --------------------------------------------------------------------------------------------------

\subsection{Programming Systems and Substrates}
\label{sec:background-sys}

\subsubsection*{Programming Systems and Substrates}
A number of systems illustrate the qualities Denicek aims to support and have a related internal
structure. Subtext, BootstrapLab and Infra~\cite{edwards-2005-subtext,jakubovic-2022-ladder,hall-2017-infra}
use structured document-based program representation and provide some of the desired programming
experiences on top of this representation. Many of those design ideas can be traced back to
Boxer~\cite{disessa-1986-boxer}, which introduced the \emph{naive realism} principle (what the user
sees is all there is) that we follow in Webnicek.

To indicate that Denicek is intended as an underlying infrastructure on top of which programming
systems can be built, we use the term \emph{computational substrate}, which also dates back to
Boxer~\cite{disessa-1995-epistemology} and is closely related to the notion of dynamic media
introduced by Kay and Goldberg \cite{kay-1977-media}. Webstrates \cite{klokmose-2015-webstrates}
is a substrate based on synchronization of documents (but without edit histories) that has been
use as the basis for multiple programming systems \cite{radle-2017-codestrates,borowski-2022-varv}.

\subsubsection*{Edit Histories and Merging}
Manipulating programs through semantically meaningful edits is a technique used by structure
editors \cite{teitelbaum-1981-cps,hempel-2018-deuce,beckman-2023-sandblocks}. The langauge of
edits has been captured formally as an \emph{edit calculus} \cite{omar-2017-hazelnut} and
edits have also been used as the basis of live programming environments \cite{storm-2013-deltas}.

Merging of edits is most frequently done in version control systems such as git. Systems based
on a more rigorous design such as Pijul \cite{meunier-2024-pijul} delay merging to a later point
by using a graph or a lattice \cite{schurmann-2022-merging}. In the context of
programming environments, Grove \cite{adams-2025-grove} uses a commutative patches, with a graph
structure similar to Pijul, as the basis for a collaborative structure editor.

More generally, merging of edits can be based on the operational transform (OT) approach
\cite{davis-2002-otsgml}, where edit conflicts are reconciled,
or conflict-free representation (CRDTs) \cite{shapiro-2011-crdts,kleppmann-2019-local,litt-2022-peritext}.
In both approaches, supporting complex edits on tree structures
remains a challenge \cite{jungnickel-2016-otjson,da-2024-jsoncrdt}. Mergeable replicated data
types (MRDTs) \cite{kaki-2019-mrdts} can merge updates provided there is a suitable relational
representation of the data. Recent work on MySubstrates and Grove \cite{adams-2025-grove,klokmose-2024-mywebstrates}
has been based on CRDTs, whereas Edwards et al. \cite{edwards-2021-typed} use OT.

% --------------------------------------------------------------------------------------------------

\begin{figure*}[t]
\includegraphics[width=0.95\textwidth,clip,trim=0cm 6cm 1.7cm 0cm]{fig/walkthrough.pdf}
\caption{Organizing a conference using Denicek. The Walkthrough shows construction of a user
  interface for adding speakers (A, B, C); refactoring of the list and merging edits (D, E); and
  formulas with schema and code co-evolution (F, G).}
\label{fig:walkthrough}
\end{figure*}

% --------------------------------------------------------------------------------------------------

\subsection{Programming Experiences}
\label{sec:background-exp}

We review a range of compelling programming experiences explored in recent research,
focusing on those supported by Denicek.

\subsubsection*{Local-First Collaboration}
Since the early collaborative programming environments such as Collabode \cite{goldman-2011-collaborative},
real-time collaboration has became widely used, if not always without challenges \cite{tan-2024-vslive}.
Merging concurrent edits is one such challenge.
In addition to CRDT-based approaches \cite{klokmose-2024-mywebstrates}, conflicts arising during
collaboration have been solved through fine-grained locking \cite{wang-2024-nbconflicts}.

\subsubsection*{Programming by Demonstration}
Earliest PbD systems used the paradigm for tasks ranging from graphics
and user interfaces to general-purpose programming \cite{smith-1975-pygmalion,cypher-1993-pbd}.
Wrangler \cite{kandel-2011-wrangler} showed the effectivity of PbD for data cleaning,
whereas more recent uses range from augmented reality prototyping \cite{leiva-2021-rapido}
and web automation \cite{chen-2023-miwa} to end-user software customizations
\cite{litt-2020-customization,litt-2020-wildcard}. Nevertheless, expressing conditions in PbD remains an
active research topic~\cite{pu-2022-semanticon,radensky-2018-conditionals}. A more general class of demonstrational
interfaces \cite{myers-2000-demonstrational} also includes programming by example, used
for example to infer data transformations in spreadsheets \cite{gulwani-2012-spreadsheets}.
Demonstrational systems can be used to directly perform actions, but also to generate code
that captures the interaction, as done e.g. in WREX \cite{drossos-2020-wrex}.

\subsubsection*{Incremental Recomputation}
Interactive programming systems with live previews \cite{petricek-2020-live,mcdirmid-2013-usable}
have been attempting to update the previews without full recomputation at least since the pioneering
work on the Cornell Program Synthesizer \cite{teitelbaum-1981-cps}. Outside live programming,
more work has been focused on updating computation when data change, although this can also
be source code of a program passed to an adaptive interpreter \cite{acar-2006-adaptive}.
Incremental recomputation is also a concern in notebook systems where cells can be run out of
order \cite{singer-2020-jollity}. This can be addressed by maintaining a dependency graph
\cite{petricek-2018-wrattler,koop-2017-dataflow} that also allows incremental recomputation
on code change.

\subsubsection*{Schema Change Control}
Schema change control is concerned with adapting data and code to reflect changes in schema.
The problem is well-studied in the context of databases \cite{brahmia-2024-evolution},
although only few systems also automatically adapt database queries \cite{wang-2019-schema}.
However, the problem is starting to be recognized in programming systems research
\cite{litt-2020-cambria,edwards-2025-schema} as well as live programming \cite{barenz-2020-live}
where state needs to be preserved during program editing.

\subsubsection*{End-User Debugging}
Non-programmers also need to be able understand and debug their programs \cite{kissinger-2006-debugging}.
Systems such as Whyline and Probe Log \cite{ko-2004-whyline,ko-2009-whyline,krebs-2023-probelog}
record information about program execution to let users analyse why they see a particular result,
whereas displaying intermediate steps can aid understanding of data science pipelines \cite{shrestha-2021-unravel}.
More generally, such functionality can leverage provenance tracking \cite{cheney-2009-provenance}
and program slicing \cite{ricciotti-2017-imperative,perera-2012-functional}. The same infrastructure
can also be used to build linked visualizations \cite{perera-2022-linked}.

\subsubsection*{Concrete Programming}
Programming can be simplified by working with concrete values instead of abstractions, an idea
pioneered by the prototype-based programming language Self \cite{ungar-1987-self}. Self uses
prototypes at the object level. At the expression level, similar functionality can be provided
by managed copy \& paste. Subtext \cite{edwards-2006-copypaste,edwards-2022-copypaste} treats this
mechanism as central, whereas other systems view tracking of copy \& paste as an additional
editor featrue \cite{jablonski-2007-cren,toomim-2004-linked}. Notably, copy \& paste has been
also tracked in spreadsheets \cite{hermans-2015-copypaste} and Gridlets \cite{joharizadeh-2020-gridlets}
offer an appealing alternative design.

\subsubsection*{Other Programming Experiences}
Several experiences not directly addressed in this paper are worth further investigation.
Projectional editors such as Lorgnette \cite{gobert-2023-lorgnette},
live literals \cite{omar-2021-livelits}, projection boxes \cite{lerner-2020-boxes} and data detectors
\cite{nardi-1998-agents} allow visualization or editing of aspects of programs through a user interface.
The problem of making live and rich editors compositional is addressed by
Engraft \cite{horowitz-2023-engraft}. Programming environments may also integrate
live examples \cite{rauch-2019-babylonian} and a variety of AI assistantce tools \cite{petricek-2023-aias,blinn-2024-llms,mcnutt-2023-nbai}.

% ==================================================================================================

\section{Walkthrough}
\label{sec:walk}

In this section, we use the web-based prototype Webnicek to illustrate programming
experiences enabled by the Denicek substrate. Webnicek is based on a structure
editor that supports navigation in the document and issuing of edit commands. In this section, we follow two
formative examples (\S\ref{app:examples}) related to conference planning.

\subsubsection*{\circled{A} Adding a Speaker}
The first user starts with an empty document, which is represented as a record. They add a field for
each heading and a field named {\small\Verb_speakers_} containing a list {\small\Verb_<ul>_}. They
use the command toolbox to add the first speaker.

\subsubsection*{\circled{B} Creating a User Interface} To simplify adding of
further speakers, the user creates a textbox and a button. They enter new speaker
details, add a new {\small\Verb_<li>_} element to the list and copy the speaker details
from the textbox into the new element using a source view.

\subsubsection*{\circled{C} Abstracting the Interaction} After adding the speaker from the textbox,
the user opens history view, selects edits that added the speaker and saves those as the
{\small\Verb_add-speaker_} interaction. They attach this interaction as a event handler for the
{\small\Verb_click_} event of the button.

\keyideabox{\faLightbulbO}{Programming by Demonstration}{Denicek implements programming by
demonstration (\S\ref{sec:impl-pbd}) by enabling the user to save and replay past interactions.
Those can be replayed directly, or used as event handlers in order to construct interactive user
interfaces (\S\ref{sec:impl-interaction}).}

\subsubsection*{\circled{D} Refactoring Document Structure} Another user starts with the initial
version of the document (A) and turns the list into a table. This is done by invoking a series of
commands that change tags, wrap elements, copy and transform values.

\subsubsection*{\circled{E} Merging Edits} The two document versions (C and D) can be automatically merged.
The refactoring is applied to all existing speakers. New speakers added using the ``Add speaker!''
button are also automatically added in the new format.

\keyideabox{\faLightbulbO}{Local-First Collaboration}{Denicek's merging
reapplies edits from another branch on top of the current history (\S\ref{sec:impl-collab}).
Merging is asymmetric, but the order does not matter in the above scenario. The same merging
operation is used when handling user interaction (\S\ref{sec:impl-interaction}).}

\subsubsection*{\circled{F} Adding Budget Calculation} A third user adds a budget calculation to
the initial document. Formulas are represented as special {\small\Verb_x-formula_} nodes whose arguments
are other nodes or references to other nodes or formulas (highlighted on mouse hover).

\keyideabox{\faLightbulbO}{Incremental Recomputation}{Formulas are document nodes and
evaluating a formula yields additional edits that augment the document with the result
(\S\ref{sec:impl-eval}). Those evaluated edits are kept at the top of the document history and are
removed in case of a conflict (\S\ref{sec:impl-incremental}), resulting in the incremental
recomputation behavior.}

\subsubsection*{\circled{G} Merging Formulas} When the budget calculation is merged with the
other edits, references in formulas are automatically updated to point to the new list.
Adding new speaker via the ``Add speaker!'' button invalidates the evaluated result.

\keyideabox{\faLightbulbO}{Schema Change Control}{The substrate understands references in
the document and updates them when applying structural edits (\S\ref{sec:impl-schema}).
Evaluation can replace formulas with values, but also augment them to enable
end-user debugging via provenance analysis (\S\ref{sec:impl-provenance}).}

% ==================================================================================================

\section{The Denicek Substrate}
\label{sec:system}
Denicek represents programs as sequences of edits that construct and transform a computational
document. In this section, we describe the structure of documents and edits, as well as the
operations that form the backbone of the system and are used to implement the
end-user programming experiences as discussed in \S\ref{sec:impl}.

% --------------------------------------------------------------------------------------------------

\subsection{Selectors, Documents and Edits}
\label{sec:system-defs}
A computational document is a tree, consisting of four kinds of nodes (Fig.~\ref{fig:doc}).
Records and lists are labeled with a \emph{tag} as in HTML. Record fields have unique names.
Lists are expected to contain elements of the same structure, identified by a unique index (serving
as an ID). Primitive nodes can be strings, numbers or references to another location in the
document tree. References can be relative or absolute. Edit operations only use absolute
references (to denote a target node), but relative references can appear in the document (e.g., to
refer to another node within a list element).

A reference is represented as a sequence of selectors (Fig.~\ref{fig:doc}).
The document model assumes that lists are homogeneous and records heterogeneous, and so the
\ident{Any} selector makes it possible to refer to all elements of a list, but there is no
way to refer to all fields of a record. Note that Denicek does not use implicit numerical
indices for lists. The index of a new list item has to be supplied explicitly. The reasoning behind
this design choice is discussed in \S\ref{sec:discuss}.

\subsubsection*{Document Edits}
The supported document edits and their behavior are listed in Fig.~\ref{fig:edits}. All edits
require a \emph{target} to which they are applied. Targets are absolute references not containing the
\ident{Parent} selector. They can contain the \ident{Any} selector, in which case the edit is applied
to multiple nodes simultaneously. Most edits can only be applied to target node(s) of a specified
kind. As discussed in \S\ref{sec:discuss}, list elements as well as fields of a record are ordered
and edits that add a new item take the index or field name of a previous node.

The edits allow an arbitrary transformation of a document via a series of steps whose effect can be
tracked by the substrate. As we saw, Denicek updates references when document structure changes.
Fig.~\ref{fig:edits} distinguishes between edits that keep references in a document unchanged
(above) and edits that affect references (below).

Renaming a field or wrapping a node updates any references to within the target location
(\S\ref{sec:system-ops}). When deleting a field to which there is a reference in the document, Denicek
rejects the edit. A copy edit of a node to which there is a reference is also rejected, because it
is ambiguous whether references referring to the original location should be left unchanged,
or modified to point to the target of the copying (a reference cannot refer to two unrelated locations).

% \note{why not have distinct Copy and Move operations to resolve this?}
% TP: Good point - this is a nice motivation for Move! I should probably add that :-)

% --------------------------------------------------------------------------------------------------

\arrayrulecolor{lightgray}
\definecolor{ekgray}{gray}{0.90}

\begin{figure}
\newcommand{\seltablecol}[3]{
\sffamily\small{\bfseries #2} & {\footnotesize #1} & \footnotesize #3\\
}
\newcommand{\ndtablecol}[4]{
\raisebox{-0.2em}{#1} & \sffamily\small{\bfseries #2}\,\;\textit{\footnotesize #3}\\[-0.2em]
&\sffamily\footnotesize #4\\[0.3em]
}

\begin{tabular}{|llp{18.08em}|}
\hline
\rowcolor{ekgray}
&&\\[-1em]
\rowcolor{ekgray}
\sffamily\small{\bfseries Selector} & {\sffamily\footnotesize Notation} & \\[0.2em]
\hline
&&\\[-1em]
\seltablecol{\Verb|..|}{Parent}{Refers to a parent of a node}
\seltablecol{\Verb|field|}{Field}{Refers to record field of a given name}
\seltablecol{\Verb|\#index|}{Index}{Refers to list element at a given index}
\seltablecol{\Verb|*|}{Any}{Refers to all children of a list node}
\hline
\end{tabular}

~\\[0.5em]

\begin{tabular}{|cl|}
\hline
\rowcolor{ekgray}
&\\[-1em]
\rowcolor{ekgray}
& \sffamily\small{\bfseries Kind}\;\,\textit{\footnotesize arguments} \\[0.2em]
\hline
&\\[-1em]
\ndtablecol{\faListUl}{List}{tag, index$_1$, child$_1$, $\ldots$, index$_n$, child$_n$}
  {Ordered list of nodes, addressable by \textit{index}. Renders as \Verb|<tag>| with children.}
\ndtablecol{\faFileO}{Record}{tag, field$_1$, child$_1$, $\ldots$, field$_n$, child$_n$}
  {Record with children addressable by \textit{field}. Renders as \Verb|<tag>| with children.}
\ndtablecol{\faExternalLink}{Reference}{selectors}
  {Reference to another document location. Displays the \Verb|selectors| as a link.}
\ndtablecol{\faFont}{Primitivie}{string \textnormal{or} number}
  {Numerical or textual primitive value. Renders as an HTML text node.}
\hline
\end{tabular}
\vspace{-0.5em}
\caption{Structure of selectors and document nodes}
\label{fig:doc}
\vspace{-1em}
\end{figure}

% --------------------------------------------------------------------------------------------------

\begin{figure}
\newcommand{\ektablecol}[5]{
\raisebox{-0.2em}{#1} & \sffamily\small{\bfseries #2}\,\;\textit{\footnotesize target, #3} & \sffamily\footnotesize #4 \\[-0.2em]
&\multicolumn{2}{l|}{\sffamily\footnotesize #5}\\[0.3em]
}
\begin{tabular}{|cp{16em}r|}
\hline
\rowcolor{ekgray}
&&\\[-1em]
\rowcolor{ekgray}
 & \sffamily\small{\bfseries Edit}\;\,\textit{\footnotesize arguments} & \sffamily\footnotesize\bfseries Target \\[0.2em]
\hline
&&\\[-1em]
\ektablecol{\faPlus}{Add}{field, after, node}{Record}
  {Add \textit{node} as a \textit{field} to the specified record \textit{after} a given field.}
\ektablecol{\faAt}{Append}{index, after, node}{List}
  {Append \textit{node} to the end of the specified list \textit{after} a given field.}
\ektablecol{\faSort}{Reorder}{permutation}{List}
  {Reorder items of a specified list according to a \textit{permutation}.}
\ektablecol{\faMinusCircle}{DeleteItem}{index}{List}
  {Delete the item at a given \textit{index} of a specified list.}
\ektablecol{\faCode}{UpdateTag}{tag}{List or Record}
  {Change the tag of a specified list or record from to a new \textit{tag}.}
\ektablecol{\faICursor}{PrimitiveEdit}{transform}{Primitive}
  {Apply primitive \textit{transform} to the specified primitive.}
\hline
&&\\[-1em]
\ektablecol{\faFont}{RenameField}{old field, new field}{Record}
  {Rename the field of a specified record from \textit{old} to \textit{new}.}
\ektablecol{\faTimesCircle}{DeleteField}{field}{Record}
  {Delete the field \textit{field} of a specified record.}
\ektablecol{\faFileO}{WrapRecord}{tag, field}{Any}
  {Wrap the specified node as a \textit{field} of a new record with \textit{tag}.}
\ektablecol{\faListUl}{WrapList}{tag, index}{Any}
  {Wrap the specified node as a sole element of a new list with \textit{tag}.}
\ektablecol{\faCopy}{Copy}{selectors}{Any}
  {Copy nodes(s) from \textit{selectors}, replacing the specified target(s). }
\hline
\end{tabular}
\vspace{-0.75em}
\caption{Summary of document edit types in Denicek}
\label{fig:edits}
\vspace{-1em}
\end{figure}

% --------------------------------------------------------------------------------------------------

\subsubsection*{Automatic Reference Update}
As discussed in \S\ref{sec:discuss}, the structure of Denicek document is implicit and not
statically enforced (the substrate can be seen as dynamically typed). As a consequence, edits
that transform the structure can be used as structural edits (affecting the structure)
or as value edits (affecting the value of specific nodes). The latter can be used, for example,
when constructing an additional list item. An item can be first constructed and then populated
with values. This temporarily violates the invariant that collections are homogeneous
(an issue we might address by using transactions as discussed in \S\ref{sec:discuss}).
Another use of value edits is when evaluating a formula, which involves wrapping the formula
node (\S\ref{sec:impl-eval}), but the operation should not affect references to the formula
itself. To support value edits, edits that normally transform references (Fig.~\ref{fig:edits} below)
have a \emph{reference behavior} option that can be set to disable automatic reference
updating. This annotation is required when the target contains the \emph{Index} selector,
which is necessarilly a value edit affecting a specific node.

% \note{I think it would be desireable to avoid all these complications. References
% should be automatically maintained so they ``just work''.}
% \note{I believe that evaluation should not change the type/shape of the document. The space
% for evaluation results should be be preallocated in some sense, either as extra structure in
% a formula, or as metadata attached to the tree. That said, I've tried various alternative
% approaches to this in Subtext without ever settling on a preferred solution.}
%
% TP: I think this is true - but that would make the system more "statically typed"
% and Denicek is dynamically typed - my only reason for this is to make the editing experience
% more "document-like" - but it is perhaps not necessary.

An important principle of the Denicek design is that the effect an edit has on references inside
the document does not depend on the current value of the document, only its type. This makes it possible to
define merging solely in terms of edits, without reference to current document state.
% (The effect \ident{Copy} has on the document structure depends on the existing document
% structure, but not on its value).
This design choice makes it impossible to encode computational logic directly in the edits
(e.g., through conditional edits). As we will see in \S\ref{sec:impl-interaction}, such logic
can be provided as an additional mechanism on top of the underlying Denicek substrate.

% --------------------------------------------------------------------------------------------------

\subsection{Primitive Operations}
\label{sec:system-ops}
Denicek provides three primitive operations. A sequence of edits can be applied to a
document, two sequences of edits can be merged and they can be checked for conflicts. Denicek
identifies edit histories by a (git-like) hash, computed from the current edit and the hash of the
preceding edit. The hash is used to identify a common shared prefix of the
history during conflict resolution and merging.

\subsubsection*{Applying Edits}
When applying an edit, Denicek locates the target node and transforms it according to the edit. If
the edit affects references in the document (Fig~\ref{fig:edits}, below), Denicek updates
the relevant references according to the rules shown in Fig.~\ref{fig:updates}, provided
that the reference behavior of the edit is not set to disable this.
%(this is required when %the target reference contains the \ident{Index} selector).

Reference update behavior for \ident{WrapRecord} and \ident{WrapList} differs in that the updated
references as a result of \ident{WrapList} use the \ident{All} selector. Although \ident{WrapList}
specifies the index to be used for the newly created list item, we assume that the operation
introduces a homogeneous list, initially containing a single element, and the updated references
should point to all eventual list items.

Note that the affected references in the document may be more specific than the edit target. For
example, if we rename {\small\Verb|old|} to {\small\Verb|new|} at {\small\Verb|/foo/*|}, a reference
{\small\Verb|/foo/3/old|} will become {\small\Verb|/foo/3/new|}. (Contrariwise, a reference in the
document cannot be more general. A more specific edit would have to contain the \ident{Index} selector
and this would require setting reference behavior to not trigger reference update.) Finally,
if the document contains a reference that would be invalidated by the \ident{Copy} or
\ident{DeleteField} edit, the edit is rejected.

% --------------------------------------------------------------------------------------------------

\begin{figure}
\newcommand{\tttablecol}[5]{
\small{\bfseries #1}\;\,\footnotesize\textit{#2}\,\; --\,\; #5\\[-0.1em]
\quad \footnotesize #3 \;\;$\Rightarrow$\;\; #4 \\[0.3em]
}
\begin{tabular}{|p{27em}|}
\hline
\\[-1em]
\tttablecol{RenameField}{target, old field, new field}{\Verb|/target/old\_field/nested|}{\Verb|/target/new\_field/nested|}
  {Replace Field for matching references.}
\tttablecol{WrapRecord}{target, tag, field}{\Verb|/target/nested|}{\Verb|/target/field/nested|}
  {Insert extra Field selector after matching prefix.}
\tttablecol{WrapList}{target, index, tag}{\Verb|/target/nested|}{\Verb|/target/*/nested|}
  {Insert extra All selector after matching prefix.}
\hline
\end{tabular}
\vspace{-0.75em}
\caption{How document edits transform references}
\label{fig:updates}
\vspace{-0.5em}
\end{figure}

% --------------------------------------------------------------------------------------------------

\begin{figure}
\newcommand{\eftablecol}[3]{
\small{\bfseries #1}\;\footnotesize\textit{target}\,\; --\,\; {\footnotesize #2}\\[-0.1em]
\quad \small #3\\[0.3em]
}
\begin{tabular}{|p{27em}|}
\hline
\\[-1em]
\eftablecol{StructureEffect}{Affects fields or structure of the target node.}
  {\ident{RenameField}, \ident{DeleteField}, \ident{WrapRecord}, \ident{WrapList}, \ident{Copy}}
\eftablecol{ValueEffect}{Transforms value, modifies list or adds an additional field.}
  {\ident{Add}, \ident{Append}, \ident{Reorder}, \ident{DeleteItem}, \ident{PrimitiveEdit}}
\eftablecol{TagEffect}{Modifies the tag of the target node.}
  {\ident{UpdateTag}}
\hline
\end{tabular}
\vspace{-0.75em}
\caption{Effects of individual edit operations}
\label{fig:effects}
\vspace{-1em}
\end{figure}

% --------------------------------------------------------------------------------------------------

\subsubsection*{Merging Edit Histories}
Merging edit histories is used when two users edit document independently, but also
when replaying edits in programming by demonstration. The operation $\mathcal{M}_E(E_1, E_2)$
works on two edit histories, $E, E_1$ and $E, E_2$, that have a shared prefix $E$.
The operation is akin to git rebase. It turns edits $E_2$ into edits $E_2'$ that can be reapplied
on top of the other edit history. Formally, $\mathcal{M}_E(E_1, E_2) = E, E_1, E_2'$.
Note that the operation is not symmetric. $\mathcal{M}_E(E_2, E_1) = E, E_2, E_1'$ and the
result of applying the two histories to the same node will differ if there are conflicts
among the edits in $E_1$ and $E_2$. We return to conflict detection in the next section.

% \note{It might help to define a function signature for merging to make this clearer: $\mathcal{M}(P, L, R) = M$ or $\mathcal{M}_P(L, R) = M$.}
% \note{Merging is often thought of as symmetric, perhaps choose a name that is clearly asymmetric, like rebase, translate, or project.}

The key operation that enables such reconciliation takes two individual edits that occurred
independently, $e_1$ and $e_2$, and produces a sequence of edits $e_2', e_2'', \ldots$ that
can be applied after $e_1$ and have the effect of $e_2$, modified to respect the effects of $e_1$.
There are two aspects of such reconciliation:

\begin{enumerate}
\item \emph{Apply to Newly Added.} If $e_2$ is adding new nodes to the document, but
  $e_1$ modified the document through a selector that would also affect the new nodes added by $e_2$,
  we need to apply the transformation represented by $e_1$ to the nodes newly added by $e_2$.
  This is done by generating an additional edit, to be applied after $e_2$, that is based on
  $e_1$ but targets only the newly added nodes (more details can be found in \S\ref{app:merge-apply-to-added}).

\item \emph{Transform Matching References.} If $e_2$ targets a node that is inside a node
  whose structure is changed by $e_1$, the target reference in $e_2$ is updated in a way that
  corresponds to the new structure. This is done using the rules shown in Fig.~\ref{fig:updates},
  that apply when transforming references inside a document, although we now also support the case
  when $e1$ is \ident{Copy} (see \S\ref{app:merge-transform-refs} for details).
\end{enumerate}

% \note{I believe that in general a ``backwards'' form of merge is also needed, as in my
% \emph{retract} and the OT \emph{Exclusion Transform}. A regular merge takes forking edits
% and composes them. The backwards form converts composed edits into forking edits (and can
% be combined with a forward merge to recompose in the opposite order). This is needed for
% cherry picking and reverts, however you don't seem to need that for this paper.}
%
% TP: Correct, I do not support those - but cherry-picking is a nice example of a thing
% that needs then and would be very nice to have!

% \note{Related question: how do you merge an Append(target, index, after, node) with
% DeleteItem(target, after)? A famous counterexample in OT was that two insertions adjacent
% to a deletion could flip their location if the edits were reordered. I'll dig up the
% details of this counterexample for you. Some systems handle this by leaving behind a
% ``tombstone'' on delete to remember the position of the deleted item. I have a novel
% solution but I'm probably going to go back to tombstones anyway because they are
% simpler and handy in the diff UI.}
%
% TP: This is a good question. My ordering of records is a somewhat last minute addition
% so it is not super well debugged. But in theory, it should be fine - because the ordering relation
% could keep the IDs of removed elements (only for ordering purposes). I think this is
% compatible with the MRDT data structure I'm trying to use.

To illustrate merging, consider a case where we created a list of work items {\small\Verb|/todo|}.
In one branch, we add an additional item to the list ($e_2$). In another branch, we wrap the
list in an extra \ident{<div>} element ($e_1$) and add a checkbox to each work item ($e_1'$):
%
\begin{equation*}
\arraycolsep=2pt
\begin{array}{rcl}
e_2 &=& \srckvd{Append}(\textnormal{\small\Verb|/todo|}, \#1, \#0, \srcid{<li>Do some work</li>})\\[0.5em]
e_1 &=& \srckvd{WrapRecord}(\textnormal{\small\Verb|/todo|}, \srcid{<div>}, \srcid{items})\\
e_1' &=& \srckvd{Add}(\textnormal{\small\Verb|/todo/items/*|}, \srcid{done}, \srckvd{nil}, \srcid{<input type="checkbox"/>})\\
\end{array}
\end{equation*}
%
If we want to append the edit $e_2$ after edits $e_1, e_1'$, we need to update its target to
reflect the wrapping (2) and we need to create and additional \ident{Add} operation
that will add the checkbox to the new item (1). The result is a sequence with two addtional edits $e_2', e_2''$:
%
\begin{equation*}
\arraycolsep=2pt
\begin{array}{rcl}
  e_1 &=& \srckvd{WrapRecord}(\textnormal{\small\Verb|/todo|}, \srcid{<div>}, \srcid{items})\\
  e_1' &=& \srckvd{Add}(\textnormal{\small\Verb|/todo/items/*|}, \srcid{done}, \srckvd{nil}, \srcid{<input type="checkbox"/>})\\
e_2'  &=& \srckvd{Append}(\textnormal{\small\Verb|/todo/items|}, \#1, \#0, \srcid{<li>Do some work</li>})\\
e_2'' &=& \srckvd{Add}(\textnormal{\small\Verb|/todo/items/\#1|}, \srcid{done}, \srckvd{nil}, \srcid{<input type="checkbox"/>})\\
\end{array}
\end{equation*}
%
If we performed the merge operation in the other order, we would append $e_1, e_1'$ after
$e_2$ without a change. In this case, the result of applying the two edit sequences would be the
same.

% --------------------------------------------------------------------------------------------------

\begin{figure}[t]
\includegraphics[width=0.95\columnwidth,clip,trim=1.7cm 6cm 7.8cm 1.5cm]{fig/merging.pdf}
\caption{Merging of two independently done sequences of edits. Two ways of merging B and C result in the same D.}
\label{fig:merging}
\end{figure}

% --------------------------------------------------------------------------------------------------

\subsubsection*{Conflict Resolution}
When merging two sequences of edits, $E, E_1$ and $E, E_2$, it is desirable that
applying $\mathcal{M}_E(E_1, E_2)$ and $\mathcal{M}_E(E_2, E_1)$ results in the same document.
This is not always the case. If the two edits modify the same value, transform the structure of
a node in incompatible ways or one edit modifies a node deleted by the other, the merge operation
cannot reconcile. Such conflicting edits can be detected, reported to the user and optionally
removed. Conflict detection is also used to implement incremental evaluation~(\S\ref{sec:impl-incremental}).

% Our treatment of conflicts necessarily differs in this respect from replication algorithms like
% Operational Transformation~\cite{OT} and Convergent Replicated DataTypes~\cite{CRDT} where
% conflicts are arbitrarily but deterministically resolved.
%
% TP: The merge operation in Denicek does also resolve them deterministically and arbitrarily
% though this does not always result in "sensible" resolution. (The dynamic nature again, I guess...)
% It is an interesting question how CRDTs could do this though - because we do want to have
% detectable conflicts (for the purpose of the evaluation)

The Denicek substrate implements a conflict detection mechanism inspired by effect
systems~\cite{lucassen-1988-effects}. The mechanism is simple and tractable, but over-approximates
conflicts, i.e.~it may report a conflict even if two edits can be merged successfully
(see \S\ref{sec:discuss}). Effects describe how an edit affects the document structure
and we distinguish between three types of effects as shown in Fig.~\ref{fig:effects}.

We say that a set of effects $F_1$ conflicts with another set of effects $F_2$ if there
are effects $f_1\in F_1$ and $f_2\in F_2$ that are of the same kind and the target
of one is a prefix of the target of the other (allowing a specific \ident{Index} to
match against \ident{All} in both directions).

Given two edit histories $E, E_1$ and $E, E_2$, Denicek can use conflict detection to remove
all conflicting edits from $E_2$ and produce a sequence of remaining edits $e_2', e_2'', \ldots$
that can be added after $E, E_1$ and do not conflict with edits in $E_1$.

To do this, we iterate over edits $e_2$ from $E_2$ and check if the dependencies of $e_2$
conflict with effects of (1) any of the effect $e_1$ from $E_1$ or (2) effects of any of the
previously removed edits. Here, the dependencies of $e_2$ include its target, but also the source
of \ident{Copy} and additional dependencies recorded explicitly as discussed in
\S\ref{sec:impl-incremental}. If a conflict is detected, the edit $e_2$ is removed and its effect
is recorded, so that we remove any subsequent edits that would depend on the removed edit.

% \note{The effect-style conflict detection is a nice idea! Removing conflicts before merging is
%  also an interesting idea I hadn't considered. That might allow me to simplify my transformation rules. }

% ==================================================================================================

\section{Programming Experiences}
\label{sec:impl}

The key claim of this paper is that the Denicek computational substrate suppports a range of
compelling user experiences, many of which crucially rely on Denicek's ability to merge edits.
We demonstrate this claim with the Webnicek web-based programming system, which uses Denicek to support:
%
\begin{enumerate}[itemsep=1.5pt,label=(\roman*),topsep=6pt,leftmargin=20pt]
\item local-first collaboration (\S\ref{sec:impl-collab}),
\item programming by demonstration (\S\ref{sec:impl-pbd} and \S\ref{sec:impl-interaction}),
\item incremental recomputation (\S\ref{sec:impl-eval} and \S\ref{sec:impl-incremental}),
\item schema change control (\S\ref{sec:impl-schema}),
\item end-user debugging via provenance tracking (\S\ref{sec:impl-provenance}),
\item concrete programming via managed copy \& paste (\S\ref{sec:impl-copy}).
\end{enumerate}
%
We describe the programming experience in isolation in the context of Webnicek in this section.
The data science notebook system Datnicek, discussed in \S\ref{sec:case}, provides
a more comprehensive evaluation by combining multiple user experiences together.

\subsection{Local-First Collaboratation}
\label{sec:impl-collab}

The Denicek substrate enables local-first collaboration~\cite{kleppmann-2019-local},
as illustrated earlier (\S\ref{sec:walk}, E). If a document is edited by multiple users, they can
each make edits to their local copy and eventually merge the variants using the operation to
merge edit histories.

Merging of histories behaves akin to git rebase in that it keeps a linear history. Synchronization
in a distributed system thus requires first reapplying local edits on top of the remote history,
before updating the remote history. Denicek thus implements the \emph{convergence model} of
document variants \cite{edwards-2025-schema}, i.e., the user cannot, for example, maintain their
own local document structure and import new data from another variant (an alternative discussed
in \S\ref{sec:discuss}).

As an example, consider the scenario in Fig.~\ref{fig:merging}. Here, document $D$ can be obtained
either by appending $C'$ (produced by the edit reconciliation operation) on top of $A,B$ or by
appending $B'$ on top of $A,C$. The edit histories resulting from the two ways of merging will
differ. In the first case ($B$ then $C'$), the \ident{Append} edit that adds a new node is followed
by further focused edits that transform the newly added node from a list item to a table row.
In the second case ($C$ then $B'$), the structural transformations are applied to all rows.

According to our effect analysis, the two operations are conflicting.
Although $B$ primarily affects the document structure, it also adds a new field to the record (email),
which is a \ident{ValueEffect}, conflicting with the \ident{ValueEffect} of $C$. In this scenario, the
conflict can be ignored and the resulting document will be the same.  However, as discussed
in \S\ref{sec:system-ops}, merging of edit histories is not symmetric and arising conflicts
can also be resolved by removing conflicting edits.

~

% \note{To guarantee that replicas converge it is necessary to have a global order on merges.
% OT and CRDT algorithms use timestamps combined with a unique ID for each client to establish
% this global order without a centralized coordinator. We can do the same. This means you never
% know if your changes will dominate or whether they will be eventually overriden but people
% seem to be willing to live with it. Personally, I'm a skeptical: I expect that like with NoSQL
% people will eventually realize this sucks and go back to centralized systems.}

% --------------------------------------------------------------------------------------------------

\begin{figure}[t]
\includegraphics[width=0.9\columnwidth,clip,trim=0cm 7cm 9cm 0cm]{fig/pbd.pdf}
\caption{Programming by demonstration is implemented by selecting edits from the document
history (1), saving them in the document (2) and replaying them (3).}
\label{fig:pbd}
\end{figure}

% --------------------------------------------------------------------------------------------------

\subsection{Programming By Demonstration}
\label{sec:impl-pbd}

In programming by demonstration \cite{cypher-1993-pbd}, the user demonstrates a task to the
system and the system then repeats it, directly or in a generalized way. To use direct
repetition with Denicek (Fig.~\ref{fig:pbd}), the user selects edits from the edit history,
name them and replay them. For general-purpose document editing in Webnicek, this requires
certain forethought, but as illustrated in \S\ref{sec:case}, the mechanism is very effective
in a domain such as data wrangling \cite{kandel-2011-wrangler}.

There are two notable aspects of our implementation of programming by demonstration in Webnicek.
First, Webnicek records the saved edits in the document itself (by representing individual edits
as nodes and storing them in a list inside the {\small\Verb_/saved-interactions_} field).
This means that no other implementation mechanism outside of the system is needed and also
that the stored edits can be modified by the user or tools working with the document.

Second, to replay edits, Webnicek does not append the recorded edits on top of the current history.
When saving edits, it stores the hash of the history at the time of saving. When replaying,
it appends the recorded edits to the top of the original history (at the time of saving) and merges
this new sequence of edits with the current history. This pushes the recorded edits through all
subsequent edits made by the user. The result can be seen in Fig.~\ref{fig:walkthrough} (E), where
a newly added speaker is transformed from a list item to a table row.

Webnicek also accounts for the case where edits recorded in the document are themselves
transformed (when they are reconciliated with other edits during merging). In this case, Webnicek
updates the recorded edits (an alternative approach is discussed in \S\ref{sec:discuss}).

% Simply replaying recorded edits provides limited programming capabilities. We discuss in the
% next section how recorded edits can themselves be edited by the user to generalize and
% conditionalize their effects. More general and conventional programming constructs are discussed in \S\ref{sec:discuss}.

% --------------------------------------------------------------------------------------------------

\subsection{Interactive User Interfaces}
\label{sec:impl-interaction}

Programming by demonstration can be used to define interactive elements in the document. In a
simple scenario, illustrated in Fig.~\ref{fig:interactive}~(1), the {\small\Verb_click_} event
handler is set to a reference to a sequence of edits recorded in the document. Clicking the button
executes the edits using the mechanism discussed in \S\ref{sec:impl-pbd}, i.e., Webnicek appends
the edits to a history at the time when the edits were recorded and merges the edits with the
current history.

The use of merging when replaying recorded edits is crucial in both the Todo App and the Conference
List formative examples (see \S\ref{app:examples}). In both cases, the merging makes it possible to define
a user interface for adding items (new Todo items, new speakers) and later change the document
structure (refactor the speaker list into a table) or add functionality (formula to evaluate whether a
Todo item has been completed) without having to recreate the user interface for adding items.
The use of merging ensures that new items are added in a correct format or with the additional
functionality.

In addition to the core functionality provided by Denicek, the Webnicek system also makes it possible
to generalize the interactions recorded through programming by demonstration in a limited way. As shown
in Fig.~\ref{fig:interactive}~(2), a button to remove all completed Todo items can be created by
generalising the {\small\Verb_remove-item_} interaction, which removes the list item at the index 0.
In addition to the recorded interaction, we manually specify (in the source view) that the edits
should be applied to all elements selected by the {\small\Verb_/items/*_} selector, instead of
the original {\small\Verb_/items/0_} selector (prefix) and that the edits should only be
applied to elements for which the formula (which tests if the checkbox is checked) specified by a
relative selector {\small\Verb_./condition/comp/result_} evaluates to true.

Note that the generalization mechanism does not violate the Denicek principle discussed in \S\ref{sec:system-defs}
that edits cannot depend on values. Webnicek finds all nodes for which the condition
holds and generates one specific (non-conditional) edit for each of the nodes.

% \note{The above paragraph was surprising to me and I'm not sure I fully understand it even after
% staring at the figure for a while. Edits are recorded as an AST using x-tags to hide them
% in the UI? That became clear only after reading the next section. Our audience might be a lot
% happier if we used a textual grammar. Earlier it was said that edits can't be conditional but
% then a conditional predicate appears here. This is a special feature of button nodes? Or event
% handler nodes? This starts to resemble Angular and similar template languages (which I find hard to read).}
%
% TP: Haha, Angualr - oh no! But I don't have a better solution right now sadly...

\keyideabox{\faMagic}{Generalization Heuristic}{Specifying generalization manually is cumbersome.
Programming by demonstration systems typically implement heuristic for generalization
\cite{myers-2000-intelligence} infers and suggests such generalizations. If integrated into
Webnicek, such heuristic could for example automatically construct a formula
based on positive and negative examples \cite{gulwani-2014-flash} (selected and deselected items).}

% --------------------------------------------------------------------------------------------------

\begin{figure}[t]
\vspace{-0.5em}
\includegraphics[width=1\columnwidth,clip,trim=0.5cm 8.5cm 8.5cm 0.5cm]{fig/interactive.pdf}
\vspace{-1.25em}
\caption{Using programming by demonstration to define a UI. The ``Add'' button (1) replays edits;
the ``Remove completed'' button (2) modifies target and specifies a condition. }
\label{fig:interactive}
\vspace{-1em}
\end{figure}

% --------------------------------------------------------------------------------------------------

\subsection{Formula Language and Evaluation}
\label{sec:impl-eval}

As illustrated earlier (\S\ref{sec:walk}, F), Denicek documents can contain formulas inspired by the
spreadsheet paradigm \cite{nardi-1990-spreadsheets}. Formulas in Webnicek do not transform the
document and their results are transient, but they can describe richer computations than what can
be expressed using document edits.

Formula evaluation leverages Denicek's ability to merge edit histories.
As illustrated in Fig.~\ref{fig:counter}, formulas are represented as document nodes with a
special tag ({\small\Verb_<x-formula>_}). They are recognized by a formula evaluator and
rendered in a special way (4), but they are created using ordinary
edits and the Denicek substrate treats them as standard nodes.

To evaluate formulas, the formula evaluator generates edits that turn the {\small\Verb_<x-formula>_} (2)
record into {\small\Verb_<x-evaluated>_} (3), which keeps the previous
formula state in the {\small\Verb_formula_} field and the evaluation result in the {\small\Verb_result_} field.
Keeping the previous state of the formula is not necessary, but it enables provenance analysis
as discussed in \S\ref{sec:impl-provenance}. The way edits generated by evaluation are merged
with the document is discussed in the next section.

% \note{As I've said before, I'd prefer to store the result of the formula inside the <x-formula>
% record instead of wrapping it on evaluation. This avoids changing the shape of the document on
% evaluation, which I recall added some complexity earlier.}
%
% TP: I think I tried that, but it still did not solve my problem - because I needed to add
% child nodes to the result. So this can probably only be done for a statically typed version.

The Counter App example shown in Fig.~\ref{fig:counter} illustrates the interaction between formulas
and programming by demonstration. To implement a counter, we first add the initial value 0 to the
document. We then perform three edits that wrap the current counter value in a formula that adds
1 to the value. The three edits are used as an event handler for the button and so clicking the
button creates an increasingly long sequence of increments. One advantage of this approach that
we will leverage later is that the system also records a trace of how the final computed value
was obtained.

% \note{OK maybe I'm finally getting it. You aren't mutating the value of the counter, you're wrapping
% the value in an increment formula. I never got that in our previous discussions. Presumably that
% evaluation is now frozen so it serves as an execution trace. This is a lot different from my
% approach where a formula stays unchanged except for the results. One advantage is that fields
% contain a record of their computed mutations whereas in my approach that is recorded in the
% past state of the doc whence the mutation occured.}

\keyideabox{\faSuperscript}{Formula Language}{
Webnicek exposes the underlying representation of formulas to the user, but the same representation
can be edited through a user-friendly mechanism such as a textual calculation view \cite{sarkar-2018-calc}
a block-based editor \cite{jansen-2019-xlblocks}. The key point
is that Denicek's tree structure makes it easy to embed formulas in document in a uniform
way, edit them as any other node and merge them with other changes.}

% --------------------------------------------------------------------------------------------------

\begin{figure}[t]
\vspace{-0.5em}
\includegraphics[width=0.95\columnwidth,clip,trim=0cm 5cm 11.5cm 0cm]{fig/counter.pdf}
\vspace{-0.5em}
\caption{We wrap the initial count (1) in a formula that adds 1 to the
value, (2). Evaluation produces the count (3), which is invalidated on subsequent clicks (4).
The button replays the saved edits that wrap the current count in a formula~(5).}
\label{fig:counter}
\vspace{-1em}
\end{figure}

% --------------------------------------------------------------------------------------------------

\subsection{Incremental Recomputation}
\label{sec:impl-incremental}

As illustrated in Fig.~\ref{fig:incremental}, Webnicek supports incremental recomputation.
The cost of speaker travel depends on the number of speakers, but the cost of refreshments
depends only on two constants. The edit that adds a speaker invalidates only the former computation.

The evaluation mechanism is illustrated in Fig.~\ref{fig:eval}. When the formulas are evaluated,
Denicek generates edits that wrap the formula in {\small\Verb_<x-evaluated>_} and set the result
to the computed value. In Webnicek, those evaluated edits are kept in a separate list that is always
appended to the top of the current edit history. When the user makes subsequent edits, the evaluated
edits are pushed through the newly added edits. If the edits conflict (according to the conflict
detection discussed in \S\ref{sec:system-ops}), the affected evaluated edits and edits that
depend on them are dropped.

The dependencies between edits alongside with the conflict detection mechanism provide functionality
that other systems implement using a dependency graph using topological sort.

\keyideabox{\faRefresh}{Live Programming}{The Webnicek prototype does not automatically
evaluate formulas. This makes it easier to understand the evaluation
mechanism, but a realistic system based on the substrate could automatically evaluate
formulas to provide a live programming experience \cite{petricek-2020-foundations,rein-2019-live}
and use incremental recomputation for performance~reasons.}

% \note{I'm confused. Changing the value of a location conflicts with an edit that evaluates a
% formula containing a reference to that location? And these evaluation edits are second-class
% edits that get retroactively dropped to ``unevaluate'' the formula? I'd prefer what you suggest
% above: formulas are seemingly globally re-evaluated after every change, with incremental
% computation a hidden perf optimization. }

% \note{An important property of incremental computation systems is avoiding redundant execution.
% Usually this involves a topological sort of the computation dependency graph. We may need to
% have a story about that to avoid hate from reviewer 2. Personally, I think incremental
% recomputation is about performance, not the programming experience, so it isn't that important for this paper. }
%
% TP: The cool thing is we get this for free! Clarified.

% --------------------------------------------------------------------------------------------------

\begin{figure}[t]
  \includegraphics[width=1\columnwidth,clip,trim=0.1cm 5cm 5.1cm 0cm]{fig/incremental.pdf}
  \vspace{-1em}
  \caption{Budget calculation based on the number of speakers (1) and the result (2). When speaker
  is added (3), only the results of affected formulas are invalidated.}
  \label{fig:incremental}
  \vspace{-1em}
\end{figure}

% --------------------------------------------------------------------------------------------------

\begin{figure}[t]
\centering
\vspace{-0.5em}
\begin{minipage}{0.55\columnwidth}
  ~\\[1em]
  \includegraphics[width=0.9\columnwidth,clip,trim=0cm 14cm 21cm 0cm]{fig/eval.pdf}
  \end{minipage}%
  \begin{minipage}{0.45\columnwidth}
    \caption{Evaluated edits (A, B) are kept at the top of the history. When they conflict with
    ordinary edits (4), they are removed (A).}
    \label{fig:eval}
  \end{minipage}
  \vspace{-1em}
\end{figure}

% --------------------------------------------------------------------------------------------------

\subsection{Schema Change Control}
\label{sec:impl-schema}

Document structure often needs to evolve \cite{burnett-2014-silos}, as illustrated by the formative
example Conference List where a list is transformed into a table. When this happens,
data and code that depend on the structure of the document need to evolve correspondingly.
The problem is well-known in database systems \cite{rahm-2006-schema} and has recently been
explored in the context of programming systems \cite{edwards-2025-schema}.

Although Denicek does not explicitly track document structure (schema or type), all documents
have an implicit structure and some edits transform this structure. As
discussed in \S\ref{sec:system-ops}, Denicek automatically updates reference nodes in the
document when edits modify the document structure. This enables a form of schema and code
co-evolution \cite{edwards-2025-schema}. Formulas embedded in documents use reference
nodes to refer to both data sources (in the document) and the results of other computations.
Consequently, if the document structure changes, the formulas are automatically updated.

Consider the example in Fig.~\ref{fig:coevolution}. The original list {\small\Verb_<ul>_} is
turned into {\small\Verb_<tbody>_} using \ident{UpdateTag} and wrapped inside {\small\Verb_<table>_}
with a field {\small\Verb_body_} using \ident{WrapRecord}. For the latter, Denicek updates the
reference accordingly turning the original {\small\Verb_/speaker_} reference in the formula
into {\small\Verb_/speaker/body_}.

Edits produced by document evaluation transform the formula structure (wrapping it in the
{\small\Verb_<x-evaluated>_} record). However, the reference behaviour of those edits is set
to keep references unchanged (as discussed in \S\ref{sec:system-defs}) and so references to
formulas are not transformed when the formula is evaluated (otherwise, references
would be updated to point to the original unevaluated formulas).

% --------------------------------------------------------------------------------------------------

\begin{figure}[t]
\vspace{-0.25em}
\includegraphics[width=1\columnwidth,clip,trim=0cm 8cm 5cm 0cm]{fig/provenance.pdf}
\vspace{-2em}
\caption{Using provenance tracking to highlight document parts that contributed to the
calculation of refreshments costs (1), travel costs (2) and all costs (3).}
\label{fig:provenance}
\vspace{-0.75em}
\end{figure}

% --------------------------------------------------------------------------------------------------

\subsection{End-User Debugging}
\label{sec:impl-provenance}

The most common kind of end-user programming question is determining whether a value they observe
is right or wrong \cite{kissinger-2006-debugging}. One way to help users answer the question is
to provide an explanation how a value was obtained \cite{ko-2009-whyline}. Webnicek provides a basic
mechanism that highlights document nodes that contributed to a specific computed result,
illustrated in Figure~\ref{fig:provenance}.

The implementation leverages the fact that evaluation keeps the original formula, as discussed
in \S\ref{sec:impl-eval}. When a formula is evaluated, the final {\small\Verb_<x-evaluated>_}
document node contains the result, but also sub-tree that represents the full evaluation trace
\cite{perera-2012-functional}. We analyse the trace, collect all reference nodes in the trace and
highlight all nodes referred to in the computation. Denicek makes this easy as we
need to collect references directly nested in the formula node.

\keyideabox{\faBarChart}{Explanations and Linked Visualizations}{In Webnicek, we implement
provenance analysis to show inputs involved in computation, but information from the execution
trace collected by the Denicek substrate can also be used to provide a detailed
explanation~\cite{perera-2012-functional} or, for example, to automatically construct
linked visualizations~\cite{perera-2022-linked}.}

% --------------------------------------------------------------------------------------------------

\subsection{Concrete Programming}
\label{sec:impl-copy}

Abstraction is an essential feature of programming, but it has a high cognitive
cost~\cite{blackwell-2002-attention}. Programming by demonstration (\S\ref{sec:impl-pbd})
offers one way of reducing the cost. Another way of making programming more
concrete~\cite{edwards-2004-example,smith-1975-pygmalion} is to support copying of functionality
as in prototype-based object-oriented programming~\cite{ungar-1987-self,randall-1995-self}.

Webnicek supports a functionality akin to managed copy \&
paste~\cite{edwards-2006-copypaste,edwards-2022-copypaste} for formulas. Rather than introducing
abstractions (functions), users can copy and modify formulas to reuse them. When the user
discovers an error in the original formula, Webnicek lets them use the merging mechanism to correct
the error in the original formula and all its copies.

In Figure~\ref{fig:copypaste}, the user copies an incorrect formula using
the \ident{Copy} edit and then modifies it to use a different data source. They then navigate back
in history to the point before the copying and create a temporary fork of the document. In the
fork, they use \ident{RenameField} to switch the argument of the division. When they merge the
temporary fork into the original document, the \emph{Apply to Newly Added} logic of the merge
operation (\S\ref{sec:system-ops}) duplicates the \ident{RenameField} edits and also applies them
to the copied formula. The merging operation with the \ident{Copy} edit and the fact that formulas
are ordinary document nodes, provide keys component for a straightforward implementation of the
managed copy \& paste functionality.

%\note{Nice and simple! My approach tried to be a lot fancier but ended up being too heavywieght. }

\keyideabox{\faClipboard}{Linked Editing}{Webnicek currently requires users to explicitly
manipulate history to correct error across multiple code clones. Research on managing duplicated
code resulted in multiple tools \cite{toomim-2004-linked,duala-ekoko-2008-clone} with dedicated
user interface to manage clones. The Denicek substrate provides the underlying mechanism that
could be used to implement a more user-friendly interface inspired by those systems.}

% --------------------------------------------------------------------------------------------------

\begin{figure}[t]
\includegraphics[width=0.9\columnwidth,clip,trim=0cm 9.5cm 8cm 0cm]{fig/coevolution.pdf}
\caption{When an edit changes the document structure, references in formulas are updated accordingly.}
\label{fig:coevolution}
\vspace{-1em}
\end{figure}

% ==================================================================================================

\section{Design Considerations}
\label{sec:discuss}

The design of the Denicek substrate is the result of an iterative process in which we repeatedly
adapted the Denicek design and revisited the implementation of the Webnicek system until we obtained
a satisfactory solution for the six formative examples (\S\ref{app:examples}). In this section, we
document the design challenges, many of which are shared with related systems
\cite{jakubovic-2022-ladder,edwards-2005-subtext,hall-2017-infra,omar-2021-livelits}.

\keyideabox{\faCubes}{Uniformity and Composability}{
Two guiding principles for the design of Denicek have been \emph{composability} and
\emph{uniformity}~\cite{jakubovic-2023-techdims}. The substrate should cover a large number of
scenarios using a small number of concepts. This is apparent in the design of the document
structure---a node can represent data, code or rich text---as well as in the role of
edits---an edit can be a value or structure change, the result of user interaction or the
result of evaluation.
}

% The inconvenience associated with composability and uniformity~\cite{jakubovic-2023-techdims} is not
% a problem for Denicek. We view Denicek as an underlying structure on top of which more
% convenient end-user programming systems can be built. The uniform design offers a degree of open-endedness,
% making it possible to use the basic structures of Denicek in new ways, not anticipated in this paper.

\subsubsection*{Dynamic and Static Typing}
Denicek does not explicitly track the document structure (it is dynamically typed). We assume that lists
are homogeneous, but do not enforce this property. A statically typed system could distinguish
structural edits from value edits. This has a theoretical appeal and it simplifies aspects
of the implementation, e.g. by removing the need to control reference updating, but it makes
working with the system less akin to document editing.

We violate structural invariants when constructing values gradually (e.g., when adding a new
speaker to a table), as well as in evaluation (wrapping {\small\Verb_<x-formula>_}
in {\small\Verb_<x-evaluated>_}). A system with static types could use a form of edit
transactions, where structural invariants are reestablished only after a sequence of edits.

% A key design choice for a substrate like Denicek is whether to track the document structure
% explicitly or make it implicit. In Denicek, we assume that elements of a list node have the same
% structure (i.e.~records should have the same fields, so that they can be targetted using the \ident{All}
% selector), but we do not enforce this. Often, the property is temporarily violated when adding a
% list item, but then restored once the item is created.
%
% Keeping the structure (type information) explicit has theoretical appeal and it simplifies some
% aspects of the implementation. In particular, edits to document structure can be clearly separated
% from edits of document values. The same separation has disadvantages in that it complicates the
% edit language and it requires user interface that departs further from document editor.
% (In Webnicek, users can select all list items and edit them at once using the same mechanism as
% when editing individual list items. With explicit structure, changing a list structure has to be
% either a different kind of edit or possibly an edit of a virtual ``prototype'' element.)

% --------------------------------------------------------------------------------------------------

\begin{figure}[t]
\includegraphics[width=1\columnwidth,clip,trim=0cm 5.5cm 1cm 0cm]{fig/copypaste.pdf}
\vspace{-1em}
\caption{Correcting an incorrect formula. The user uses Copy to create the second
formula (1). They notice an error (2), go back in history to switch left and right argument (3),
merge the change and re-evaluate both formulas (4).}
\label{fig:copypaste}
\vspace{-0.5em}
\end{figure}

% --------------------------------------------------------------------------------------------------

\subsubsection*{Expressive Selectors and Edits}
Our merge operation operates on two sequences of edits, but it does not need access
to the current document state or history. This limits the expressiveness of edits. Denicek does not
support conditional edits (applied only when a specific condition holds), because merging would then
not be able to determine if two edits are conflicting. As a result, generalization of recorded
interactions (\S\ref{sec:impl-interaction}) has to be done at a meta level and generates a series
of individual edits. This design also limits the expressiveness of selectors. For example,
an edit targeting nodes with a specific tag (e.g., all {\small\Verb_<h3>_} elements) would also be
conditional.

% One specific design choice in Denicek is that the merge operation operates solely on two
% sequences of edits, but it does not need access to the current document state (or full edit
% histories). The reconciliation of edits (\S\ref{sec:system-ops}) is defined for two individual
% edits $e_1$ and $e_2$.
%
% This design limits the possible language of edits. We cannot, in general, support
% conditional edits (that are applied only when a certain condition holds for
% the current document state). This is because the merging operation would not be able to
% determine whether the edit is applied (and whether its effects should~take~place).
%
% One consequence of the design choice is that the generalization of recorded interactions (\S\ref{sec:impl-interaction}) has to
% be done on a meta-level (by specifying how to invoke edits, rather than through the edits themselves).
% Another consequence is that our langauge of selectors has to be abstract.
% We can target all list elements using the \ident{All} selector, but cannot combine this
% representation with a representation that explicitly lists indices of all items currently in the list.

% Denicek supports only a limited set of selectors. An absolute reference can select a record
% field (\ident{Field}), specific list item (\ident{Index}) or all list items (\ident{All}).
% Once can imagine a range of other useful selectors, for example to select list items with a
% given tag or select items satisfying other conditions.
%
% We could support such selectors for those edits that do not affect the document
% structure. However, supporting such conditional selectors in general is incompatible with our
% design decision that merging should not depend on document state. (If an edit wraps the body of all
% {\small\Verb_<h3>_} elements, we need to know the document state in order to determine whether
% we need to modify the selector pointing to an item at a specific index.)

\subsubsection*{List Indices and Ordering}
Denicek does not use numerical indices for lists. Indices are unique identifiers that have to
be provided explicitly. (Although Webnicek generates indices automatically when editing lists.)
The design supports a programming by demonstration scenario where the user adds a new list item
and then modifies it (adding a speaker or a Todo item).

With numerical indices, the edits following \ident{Append} would not have access to the index of
the newly added item. Using a numerical index computed from the list length would require
complex logic to update indices when merging edit operations that affect lists. Another alternative
is to disallow modification of a newly added item, but this limits programming by demonstration.

% Denicek does not use implicit numerical indices for lists. When adding an item, the
% user (or the system built on top of Denicek) has to explicitly provide an index. This design
% is motivated by a typical programming by demonstration scenario where the user adds a
% new list item and then modifies it (adding a speaker or a Todo item). When using explicit indices,
% the edits following \ident{Append} can use the index to refer to and modify the newly added
% item. (To avoid overwriting existing items when replaying recorded edits, Webnicek replaces
% indices of newly added items when replaying recorded interactions in \S\ref{sec:impl-pbd}). \note{I don't think that was explained back there.}
%
% Using implicit numerical indices is possible, but it requires treating all list-related edits
% as operations that transform document selectors. This means adding multiple new rules to
% Figure~\ref{fig:updates}, which specifies how edits transform references. Merging then has to
% modify indices when edits add, remove or reorder list items. List order then, somewhat confusingly,
% becomes a part of document structure rather than a property of the list value.
%
% \note{I'm still unclear whether your list indices are ordinal numbers like in an array or are unique IDs. Your usage of implicit/explicit isn't obvious to me. The above paragraph seems to imply they are IDs. Note that automatically generating unique IDs in code execution can lead to nondeterminism. If you are using IDs, do they affect equality of lists?}
% The problem could also be avoided by not letting the user modify a list item after adding it.
% (We can require new items to be constructed outside of the list and then added through a
% single edit that combines the logic of \ident{Append} and \ident{Copy}.) This is easy to
% implement, but it requires unintuitive user interaction and it also does not address all
% use cases for modifying list item via index. \note{I'm still confused, but this sounds like it really isn't a solution afterall.}

To maintain order of list elements, Denicek uses a data structure inspired by the list MRDT
\cite{kaki-2019-mrdts} and requires specifying the index of a preceding item when appending
or inserting into a list. Denicek uses the same mechanism for record fields as the order of fields
may be visible to the user, for example in Webnicek where fields represent children of a HTML node.
Making the order of fields explicit ensures that it is maintained during merging.

% \subsubsection*{Ordering List Items and Record Fields}
% When list indices are not numerical and supplied explicitly, we need another mechanism for
% tracking order of list items. Denicek uses a data structure inspired by Mergeable Replicated
% Data Types \cite{kaki-2019-mrdts} and requires specifying the index of a preceding list item
% when appending to a list. However, Webnicek supplies the index of preceding item automatically.
% (If two edits append to the same list independently with the same preceding item, the order
% is non-deterministic.) \note{Why is this needed? I let the order of operations determine the list order. }
%
% Note that Denicek uses the same mechanism for record fields. Keeping record fields ordered is
% desirable because Denicek documents map directly to HTML documents and so the order is visible
% to a user. Moreover, different ways of merging edits can result in different order of adding fields
% to a record. Making order explicit ensures those result in equivalent documents.
%
% Denicek still makes a distinction between records and lists, even if the two are similar
% technically. The difference is conceptual. A list is expected to contain items of the
% same structure (that can be addressed using the \ident{All} selector), whereas record fields
% are expected to be of different types. If we unified lists and records, \ident{All} would have
% to be applicable to records too and inferring and checking document structure would be
% challenging.

% TODO: This is potentially interesting too, but it is maybe way too complex and weird
%
% \subsubsection*{Dependency Tracking}
% Recall that the incremental recomputation (\S\ref{sec:impl-incremental}) is based on the
% conflict detection mechanism of Denicek (\S\ref{sec:system-ops}). When pushing edit $e_2$ through
% $e_1$, a conflict is detected if they affect overlapping targets or if $e_1$ modifies a
% document node that is a source of $e_2$, typically, the source of a \ident{Copy} edit.
%
% For edits generated during formula evaluation (\S\ref{sec:impl-eval}), we need to track an
% additional kind of dependency. Evaluating a formula that adds two numbers from two document
% location results in an edit that sets the result of the formula to the sum of the two numbers.
% The edit contains the resulting number, but we also need to record that the edit has the two
% source locations as dependencies. For this reason, Denicek makes it possible to annotate edits
% with additional dependencies. One alternative would be to introduce a special \ident{Evaluate}
% edit that would identify an operation to compute and a list of arguments, specified as a list
% of references. We choose the former to keep the set of edits smaller.

\subsubsection*{Merging Structure and Capabilities}
Denicek keeps a linear history and merging appends edits to the top of the history.
This model is akin to git rebase. One alternative is to maintain a graph of edits akin
to git merge. This would simplify recording of edits in programming by demonstration
(\S\ref{sec:impl-pbd}) as we would not need to update saved edits transformed
during merging. (Their hashes would remain the same.) However, supporting special
merge edits and non-linear history would make the basic substrate more complex.

Denicek also supports only the simpler \emph{convergence} model of collaborative editing \cite{edwards-2025-schema}
where users have to merge all changes in order and cannot merge only selected edits (cherry
picking in git). The \emph{divergence} model lets users, for example, keep their own schema but
import all data edits. Supporting the model requires a \emph{retract} operation \cite{edwards-2021-typed}
that is dual to our edit reconcilliation (given two subsequent edits, $e_1, e_2$, we want
to generate $e_2'$ that has the same effect as $e_2$ but can occur before $e_1$).

\section{Case study}
\label{sec:case}
(data science environment)

what are the limitations of Datnicek

x

x

x

x

\section{Evaluation}
\label{sec:eval}
how well did Denicek work for implementing Datnicek

NICE THINGS IN DATNICEK
* merging of code edits (e.g. change parameter merges with adding a call to chain)
* remove cell to which there are refs - does the susbtrate catch this?

\section{Discussion}

\bibliographystyle{plain}
\bibliography{paper}

\newpage
~
\newpage

\appendix
\section{Formative Examples}
\label{app:examples}

The design the Denicek substrate, we identified six formative examples shown in
Fig.~\ref{fig:examples}. The examples range from established industry benchmarks
(Todo and Counter apps) to cases from literature~\cite{edwards-2022-copypaste} and
problems posed as schema change challenges \cite{edwards-2025-schema}.
The Denicek substrate then co-evolved with Webnicek, a simple web-based programming environment
built (as directly as possible) on top of the substrate and was used to solve implement the
formative examples.

Many of the formative examples include a small programming challenge, such as adding
user interface to add a new speaker, a new list item or modify the count. Our aim was for
the substrate to enable solving  those through programming by demonstration.
Programming by Demonstration is often used in data wrangling
\cite{gulwani-2012-spreadsheet,drossos-2020-wrex,gulwani-2014-flash}. Our Hello World example
is only a minimalistic illustration of such use, loosely inspired by earlier work \cite{miller-2001-simult}.

\section{Merging Edit Histories}
Recall that merging takes two edit histories, $E, E_1$ and $E, E_2$,
transforms edits $E_2$ into $E_2'$ that can be reapplied on top of the first history
resulting in $E, E_1, E_2'$. The key operation takes two individual edits, $e_1$ and $e_2$
and produces a sequence of edits $e_2', e_2'', \ldots$ that can be applied after $e_1$,
and combine the two edits. This section provides details about the two aspects of this operation.

\subsection{Apply to Newly Added}
\label{app:merge-apply-to-added}

Assume that edits $e_1$ and $e_2$ occurred independently. We want to modify $e_2$ so that
it can be placed after $e_1$. If the edit $e_2$ added new nodes to the document that the
edit $e_1$ would affect, we generate an additional edit that apply the transformation of $e_1$
to the newly added nodes (and only to those).

The only edits that add new document nodes are \ident{Add}, \ident{Append}, \ident{Copy} and
so we consider this case if the edit $e_2$ is one of those. If so, we check
whether the target of $e_1$ is within the target of $e_2$, i.e., the list of selectors that
forms the target of $e_2$ is a prefix of the list of selectors that forms the target of $e_1$.

Along the way, we compute a \emph{more specific prefix}. If the target of $e_1$ contains
the \ident{All} selector, it can be matched against a specific \ident{Index} selector in the
target of $e_2$ (if the selector of $e_1$ is more specific than that of $e_2$, the targets
are not matched). We then replace the original prefix in $e_1$ with the \emph{more specific prefix}
that contains \ident{Index} selector in places where the original edit contained \ident{All}.
This way, we obtain $e_1'$ which is a focused version of $e_1$ that applies only to the
nodes newly added by $e_2$. The edit $e_2$ thus becomes a pair of edits $e_2, e_1'$.
The final document will contain edits $e_1, e_2, e_1'$ -- that is, it will first apply the edit
$e_1$ to nodes already in the document, then add new nodes and then apply the transformation
represented by $e_1$ to the newly added nodes.

\subsection{Transform Matching References}
\label{app:merge-transform-refs}

As above, assume that edits $e_1$ and $e_2$ occurred independently. We want to
modify $e_2$ so that it can be placed after $e_1$. If $e_1$ is any of the three edits
listed in Fig.~\ref{fig:updates} (\ident{RenameField}, \ident{WrapRecord}, \ident{WrapList}),
we collect all references that appear inside $e_2$ (the target, the source of \ident{Copy}
and any references occurring in the nodes added by \ident{Add} or \ident{Append}).
If the target of $e_1$ is a prefix of any of those references, we update the references
accordingly and obtain a new edit $e_2'$. Note that it would be an error to match specific
\ident{Index} in $e_1$ with more general \ident{All} in $e_2$, but this cannot happen --
reference updating is not done when the target of $e_1$ contains \ident{Index}.

Now consider the case when $e_1$ is \ident{Copy} and the edit $e_2$ targets a node that is
the source node of the copy operation (or any of its children). In this case, it is reasonable
to require that the edit $e_2$ is applied to both the source and the target of the copy.
(This is required by the refactoring done in the Conference Budget example.)
We handle the case by creating a copy of $e_2$ with transformed selectors (target and, if $e_2$
is also \ident{Copy}, also its source). To transform the selectors, we replace the prefix
formed by the source of the \ident{Copy} by a new prefix, formed by the target target of
the \ident{Copy}. We then add the new operation as $e_2'$ if at least one of its selectors
was transformed (typically target, but possibly also source).

\begin{figure}
\newcommand{\extablecol}[5]{
\small{\bfseries #1} #2\,\; --\,\; #3\\[-0.1em]
\quad \footnotesize #4 \\[-0.2em]
\quad \footnotesize \emph{#5} \\[0.4em]
}

\begin{tabular}{|l|}
\hline
\\[-1em]
\extablecol{Counter App}{\cite{kiss-2014-7guis}}
  {Counter with increment and decrement buttons.}
  {The current count is represented by a formula that is modified by the buttons.\\[-0.2em]\quad\footnotesize
  The user can inspect the evaluation trace to see how the count was modified.}
  {Programming by Demonstration, Incremental Recomputation, End-User Debugging}
\extablecol{Todo App}{\cite{osmani-2024-todomvc}}
  {Buttons to add an item and remove all completed.}
  {Adding an item must correctly merge with independently added functionality to \\[-0.2em]\quad\footnotesize
   compute which items are completed and remove them based on a formula result.}
  {Programming by Demonstration, Local-First Collaboration, Incremental Recomputation}
\extablecol{Conference List}{\cite{edwards-2025-schema}}
  {Manage a list of invited conference speakers.}
  {Adding speakers to a list through an in-document user interface merges with \\[-0.2em]\quad\footnotesize
   refactoring that turns the list into a table and separates name from an email.}
  {Local-First Collaboration, Programming by Demonstration}
\extablecol{Conference Budget}{\cite{edwards-2025-schema}}
  {Calculate budget based on a speaker list.}
  {References are updated  when the list is refactored. Only affected formulas\\[-0.2em]\quad\footnotesize
   are recomputed and the user can view elements on which the result depends.}
  {Local-First Collaboration, Incremental Recomputation, End-User Debugging}
\extablecol{Hello World}{\cite{miller-2001-simult}}
  {Normalize the capitalization of two word messages.}
  {An operation to normalize the text in a list item can be recorded and applied to\\[-0.2em]\quad\footnotesize
   all list items or, alternatively, applied directly to all list items.}
  {Programming by Demonstration}
\extablecol{Traffic Accidents}{\cite{edwards-2022-copypaste}}
  {Compute statistics using two data sources.}
  {Formula to compute statistics can be reused with a different data source;\\[-0.2em]\quad\footnotesize
   error correction is propagated automatically to the copied version of the formula.}
  {Concrete Programming, Incremental Recomputation}
\hline
\end{tabular}

\caption{Formative examples used in Denicek design}
\label{fig:examples}
\vspace{-1em}
\end{figure}

\end{document}